\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-09-11 analysis/manuscript.Rmd
Reviewer 1 suggestions to improve the introduction

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted in transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from the mid-2010s onwards, where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
Similar to big deals, transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, but aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023]. -->

@@

By the end of 2023, many transformative agreements had been implemented, but the interim outcomes were mixed.
The ESAC Transformative Agreement Registry[^esacregistry], the largest resource for library consortia to publicly share their agreements, recorded more than 800 transformative agreements.
These agreements resulted in the publication of up to 900.000 open access articles published in both fully open access and hybrid journals, according to the accompanying ESAC Market Watch[^esacmarketwatch]. 
Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilisation of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023]. 
The ongoing standardisation of transformative agreements has contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuates market concentration [@Butler_2023; @Shu_2023]. 
Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
The focus on large commercial publishers might also increase inequality [@Ross_Hellauer_2022], because transformative agreements' focus on pay to publish mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest]. 
Besides, an editorial-board resignation raised concerns that transformative agreements might encourage publishers to maximize journal publication volume "without regard to quality" [@Goodin_2023].

>>>>

As a consequence, libraries and their consortia began to develop licensing strategies aimed at avoiding such ‘double dipping’ scenarios, in which well-established commercial publishers gain twice from reading and open access publishing fees [@Mittermaier_2015], as well as to increase publisher-provided immediate open access [@BJ_RK_2014; @Schimmer_2015].
These considerations resulted in transformative agreements[^agreements], which cover a broad range of contracts between library consortia and publishers from the mid-2010s onwards, where institutional spending for subscriptions and open access publishing are considered together [@Borrego_2020; @Hinchliffe_2019].
Transformative agreements seek to control costs while allowing a transitional phase for publishing more open access articles. 
Similar to big deals, transformative agreements mainly bundle hybrid and subscription-only journals from commercial publishers, but aim at a higher degree of transparency than previous big deals, where contracts including payments were confidential [@Bergstrom_2014].
<!-- For example, the nationwide DEAL consortium in Germany, which includes universities and academic institutions, collectively cancelled their subscriptions with Elsevier from 2017 onwards to negotiate a transformative agreement [@Fraser_2023]. -->

@@

By the end of 2023, many transformative agreements had been implemented, but the interim outcomes were mixed.
The ESAC Transformative Agreement Registry[^esacregistry], the largest resource for library consortia to voluntarily and publicly share their agreements, recorded more than 800 transformative agreements.
These agreements resulted in the publication of up to 900.000 open access articles published in both fully open access and hybrid journals, according to the accompanying ESAC Market Watch[^esacmarketwatch]. 
Library consortia reported increased open access volume, streamlined payment and monitoring procedures, as well as extensive utilisation of open access options by the researchers they serve [@Marques_2020; @Pinhasi_2020; @Parmhed_2023]. 
The ongoing standardisation of transformative agreements has contributed to improved transparency in terms of contracts and publisher-provided article metadata [@Marques_2019; @Pinhasi_2021].
However, with the growing trend toward transformative agreements, continued reliance on big deals is perceived as problematic, because it perpetuates market concentration [@Butler_2023; @Shu_2023]. 
Whether transformative agreements lead to reduced pricing remains uncertain [@Borrego_2023] and
a substantial transition of hybrid journals towards full open access could not be observed [@Momeni_2021; @Matthias_2019].
The focus on large commercial publishers might also increase inequality [@Ross_Hellauer_2022], because transformative agreements' focus on pay to publish mainly targets institutions from high-income countries, furthering a questionable journal prestige culture [@budapest]. 
Besides, an editorial-board resignation raised concerns that transformative agreements might encourage publishers to maximize journal publication volume "without regard to quality" [@Goodin_2023].
